\def\filepath{templates}
%\def\filepath{C:/Users/holden-lee/Dropbox/Math/templates}

\input{\filepath/packages_book.tex}
\input{\filepath/theorems_with_boxes.tex}
\input{\filepath/macros.tex}
\input{\filepath/formatting.tex}
\input{\filepath/other.tex}
\input{\filepath/theorem_num.tex}

\def\name{Metric embeddings and geometric inequalities}


\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
\lhead{MAT529}
\chead{} 
\rhead{Metric embeddings and geometric inequalities} 
\lfoot{} 
\cfoot{\thepage} 
\rfoot{} % !! Remember to change the problem set number
\renewcommand{\headrulewidth}{.3pt} 
%\renewcommand{\footrulewidth}{.3pt}
\setlength\voffset{0in}
%\setlength\textheight{648pt}

\begin{document}
%\input{\filepath/titlepage.tex}
%\maketitle
\title{Metric embeddings and geometric inequalities}
\author{Lectures by Professor Assaf Naor \\ Scribes: Holden Lee and Kiran Vodrahalli}
\maketitle


\tableofcontents

%\startcontents
%\printcontents{ }{-1}{}

\chapter*{Introduction}

%Princeton, Spring 2016. 

The topic of this course is geometric inequalities with applications to metric embeddings; and we are actually going to do things more general than metric embeddings: Metric geometry. Today I will roughly explain what I want to cover and hopefully start proving a first major theorem. 
The strategy for this course is to teach novel work. Sometimes topics will be covered in textbooks, but a lot of these things will be a few weeks old. There are also some things which may not have been even written yet. I want to give you a taste for what's going on in the field. Notes from the course I taught last spring are also available.

One of my main guidelines in choosing topics will be topics that have many accessible open questions. I will mention open questions as we go along. I'm going to really choose topics that have proofs which are entirely self-contained. I'm trying to assume nothing. My intention is to make it completely clear and there should be no steps you don't understand. 

Now this is a huge area. I will explain some of the background today. I'm going to use proofs of some major theorems as excuses to see the lemmas that go into the proofs. Some of the theorems are very famous and major, and you're going to see some improvements, but along the way, we will see some lemmas which are \emph{immensely powerful}. So we will always be proving a concrete theorem. But actually somewhere along the way, there are lemmas which have wide applicability to many many areas. These are excuses to discuss methods, though the theorems are important. 

The course can go in many directions: If some of you have some interests, we can always change the direction of the course, so express your interests as we go along. 

%Notes from Assaf Naor's class ``Geometric inequalities with applications to metric embeddings" at 

%More generally we'll cover metric geometry.

%I'll teach things that just appeared or are still in the works. Most of them aren't in books yet; some of them haven't even be written yet. I'll give a taste of what's going on in the field. I'll choose topics with many open questions associated with them and with self-contained proofs. 

%This is a huge area. I'll use proofs of major theorems as excuses to see the lemmas that go into the proofs. The theorems are famous; in addition to being famous, they use lemmas that are immensely powerful. Instead of presenting the geometric inequalities and then applications afterwards, we'll see the lemmas on the way to proving these big theorems.

%twists in proofs new, useful.
%Express your interest---I'm flexible.

%I taught a course.
%no overlap.

%some commericals on what to prove later, puzzles.

%\setcounter{chapter}{1}
\input{2-1.tex}
\input{2-3.tex}
\input{2-8.tex}
\input{2-10.tex}
\input{2-15.tex}
\input{2-17.tex}
\input{2-22.tex}
\input{2-24.tex}
\input{2-29.tex}
\input{3-2.tex}
\input{3-7.tex}
\input{3-21.tex}
\input{3-23.tex}
\input{3-28.tex}
\input{3-30.tex}
\input{4-11.tex}
%INSERT_HERE

%\appendix
%\input{distribution_chapters/a.tex}

%\bibliographystyle{plain}
%\bibliography{\filepath/refs}

\printnomenclature
\printindex
\end{document}